---
title: "Kaggle Axa Project"
author: "Alejandra Jaimes, Alex Singal, Christopher Peter Makris, Dani Ismail"
date: "July 2015"
output: html_document
---



```{r echo=TRUE, results='hide',message=FALSE}
library(ggplot2)
library(GGally)
library(corrplot)
```


###Loading Data
```{r}
load("Data/small_axa.RData")
```

### Data preparation
Preparing data frame by converting variables to numeric. 
```{r}
dim(axa) #600 Observations, 88 Variables
names(axa)
sapply(axa, class)
table(sapply(axa, class)) #Characters need to be converted to numeric type.

###Converting character variables to numeric variables.
converted.columns = sapply(axa[, 2:87], as.numeric)
axa.converted = data.frame(axa[, 1], converted.columns, axa[, 88])
names(axa.converted) = names(axa) 

axa.converted$driver = as.factor(substr(axa.converted$driver_trip, 1, 1))
names <- names(axa.converted)
names <- gsub("-", "_", names)
names(axa.converted) <- names
remove(axa)
axa = axa.converted
remove(axa.converted)
```

### Data Exploration

```{r}
summary(axa)
```


Checking for NA values. 
```{r}
sum(is.na(axa)) #3,375 missing values.
na.list = as.data.frame(sapply(axa, function(x) { sum(is.na(x)) })) #Focused in certain variables.
na.list$variable = rownames(na.list)
colnames(na.list) = c("sum.na","variable")
rownames(na.list) <- NULL
na.list[order(na.list$sum.na, decreasing = TRUE), ]
```


Creating boxplots split by driver type and response.
```{r}

var_list = names(axa)[1:89]
boxplot_list = list("null")

for (i in 2:88) {
    p = ggplot(data=axa,aes_string(x=var_list[[89]], y=var_list[[i]],
                                   fill= var_list[[88]])) + 
          geom_boxplot() + ggtitle(i)                   
    boxplot_list[[i]] = p
    }
boxplot_list [[79]]
```

Creating plot
```{r, eval = FALSE} 
#change eval = TRUE to create this plot. It requires a lot of computation.
axa2 = axa
axa2$isDriver = as.factor(axa2$isDriver)
ggpairs(data=axa2[,c(3, 4, 5, 7, 8, 18, 23, 24, 28, 32, 34, 48, 74, 75, 76, 77, 78, 79, 84, 85,88)], color ="isDriver" )
```


From the boxplots, variable `r  colnames(axa[79])` show high variances when comparing different drivers and the reponse. 

However this variable has 552/600 empty values. Since we consider this can be a variable that could predict the reponse. We decided to impute ths variable and check if we can use it in our model. 

Checking the correlation with variable 79 (`r  colnames(axa[79])`) and selecting the max and min correlations. 
```{r}
correlation_axa = cor(axa[,79] , axa[,2:87] , use = "pairwise.complete.obs")
cor_axa_t = t(correlation_axa)
which.min(cor_axa_t[-78,])
cor_axa_t[2,]
which.max(cor_axa_t[-78,])
cor_axa_t[3,]

```
The variables with the highest correaltions are: `accel_quantile.0.` and `accel_quantile.100.`. 

We picked variable `accel_quantile.0.` with the highest correlation with variable 79.

Using linear regression to impute the values. 

First model using variable accel.quantile.0 variable to predict attr.hv.sd 
```{r}
model = lm(attr.hv.sd ~ accel_quantile.0., data = axa)
summary(model)
par(mfrow=c(2, 2))
plot(model)

par(mfrow=c(1, 1))
plot(axa$accel_quantile.0., axa$attr.hv.sd)
abline(model, col = "red3", lwd = 2)
```
From the fitted vs residuals values plot it can be seen that the plot shows a pattern. 

Checking for transformations:

```{r}
library(MASS)
bc = boxcox(model)
which(bc$y == max(bc$y))
bc$x[54]
```
The boxcox plot shows that we can apply a transformation to remove the pattern. 

Transforming the data taking the max point of the boxcox transformation, and using it for a new model:

```{r}
transformed = axa$attr.hv.sd^bc$x[54]
hist(transformed, prob = TRUE)
lines(density(transformed, na.rm = TRUE), col = "red3", lwd = 2)
qqnorm(transformed)
qqline(transformed)
shapiro.test(transformed)

model2 = lm(transformed ~ accel_quantile.0., data = axa)

par(mfrow=c(2, 2))
plot(model2)

par(mfrow=c(1, 1))
col.vec = rep(1, nrow(axa))
col.vec[305] = "red3"
plot(transformed ~ axa$accel_quantile.0., col = col.vec, pch = 16)
abline(model2, col = "red3", lwd = 2)
```
From model2, a leverage point (observation 305) is analyzed.

In the following model the levarage point is removed from the imputing model.


```{r}
acceleration = axa$accel_quantile.0.[-305]


model3 = lm(transformed[-305] ~ acceleration, data = axa)
par(mfrow=c(2, 2))
plot(model3)

par(mfrow=c(1, 1))
plot(acceleration, transformed[-305])
abline(model3, col = "red3", lwd = 2)
abline(model2, col = "blue", lwd = 2, lty = 2)

summary(model3)
```
The R-squared value after removing the levarage improved from `r summary(model2)$adj.r.squared` to `r summary(model3)$adj.r.squared`

Model3 is now used to impute the missing values.

```{r}
input = axa$accel_quantile.0.[is.na(axa$attr.hv.sd)] #Selecting predictor values with attr.hv.sd = NA. 

index = which(is.na(axa$attr.hv.sd)) 
imputed.values = predict(model3,  newdata = data.frame(acceleration=input))

axa$attr.hv.sd.imputed = axa$attr.hv.sd^bc$x[54]
axa[index,"attr.hv.sd.imputed"]  = imputed.values
```


``` {r}
base.logistic = glm(isDriver ~ attr.hv.sd.imputed + driver, family = "binomial", data = axa)
summary(base.logistic)

ggplot(data = axa, aes_string(x = "driver",
                              y = "attr.hv.sd.imputed",
                              fill = "isDriver")) +
  geom_boxplot()

ggplot(data = axa, aes_string(x = "driver",
                              y = "attr.hv.sd.imputed",
                              fill = "isDriver")) +
  geom_boxplot() +
  ylim(1, 1.1)

length(which(base.logistic$fitted.values >= .5))

library(VIM)
axa.knn = kNN(axa)
axa.imputed = axa.knn[, 1:90]

axa.numbers = axa.imputed[,-c(1, 88, 89)]

pca = prcomp(axa.numbers, center = TRUE, scale = TRUE)
summary(pca)

plot(pca, npcs = 87, type = "lines", ylim = c(0, 15))

variances = pca$sdev^2
total.variance = sum(variances)
prop.variance = variances/total.variance
plot(prop.variance)

par(mfrow = c(2, 5))
for (i in 1:10) {
  plot(pca$rotation[, i])
abline(h = 0, lty = 2)
}

test = glm(isDriver ~ . - driver_trip, family = "binomial", data = axa.imputed)
summary(test)



axa.new = axa[, -c(1, 74:79, 90)]
axa.knn = kNN(axa.new)
axa.imputed = axa.knn[, 1:82]

base = glm(isDriver ~ 1, data = axa.imputed)
saturated = lm(isDriver ~ ., data = axa.imputed)
scope = list(lower = formula(base), upper = formula(saturated))

forward = step(base, scope, direction = "forward")
backward = step(saturated, scope, direction = "backward")
both.base = step(base, scope, direction = "both")
both.saturated = step(saturated, scope, direction = "both")

AIC(forward, backward, both.base, both.saturated)
BIC(forward, backward, both.base, both.saturated)

summary(both.base)
library(car)
vif(both.base)

manual.model = glm(isDriver ~ jerk_quantile.25. + accel_quantile.25. + angle_mean + angle_change + jerk_quantile.100. + attr.turn.ap.mean + attr.turn.ap.sd + angle_quantile.100. + direct_dist + dist + jerk_quantile.50., family = "binomial", data = axa.imputed)


pca = prcomp(axa.imputed[,-c(81, 82)], center = TRUE, scale = TRUE)
summary(pca)

plot(pca, npcs = 87, type = "lines", ylim = c(0, 15))

variances = pca$sdev^2
total.variance = sum(variances)
prop.variance = variances/total.variance
plot(prop.variance)

par(mfrow = c(2, 5))
for (i in 1:10) {
  plot(pca$rotation[, i])
abline(h = 0, lty = 2)
}

pca1 = pca$x[, 1]
pca2 = pca$x[, 2]
pca3 = pca$x[, 3]
pca4 = pca$x[, 4]
pca5 = pca$x[, 5]

pca.model = glm(axa.imputed$isDriver ~ pca1 + pca2 + pca3 + pca4 + pca5, family = "binomial")

test = glm(isDriver ~ attr.hv.sd + driver, family = "binomial", data = axa)

```

```{r}
library(randomForest)

model = randomForest(as.factor(isDriver) ~ .,
                     ntree = 1000,
                     data = axa.imputed)
model

output = predict(model, axa.imputed[, -81], type = "response")
output

table(output, axa.imputed$isDriver)

library(caret)
ctrl = trainControl(method = "repeatedcv",
                     number = 10, repeats = 10)
grid_rf = expand.grid(.mtry = 10)
m_rf = train(as.factor(isDriver) ~ .,
             data = axa.imputed,
             method = "rf",
             metric = "rocc", #####Kappa???
             trControl = ctrl,
             tuneGrid = grid_rf)
m_rf

```
